---
title: "M4 Competition"
author: "VuThanh"
date: "March 21, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Import packages
```{r, message=FALSE, warning=FALSE}
library(M4comp2018)
library(xts)
library(astsa)
library(ggplot2)
library(forecast)
library(ggfortify)
library(fpp2)
```
## 2. Load M4 data
```{r, message=FALSE, warning=FALSE}
data(M4)
```

## 3. Create a data frame to sumarize the M4 data structure
```{r, message=FALSE, warning=FALSE}
df = data.frame(matrix(ncol = 5, nrow = 100000))
colnames(df) = c("st", "n", "type", "h", "period")
df$st = unlist(Map(function(l) {as.character(l$st[[1]][1])}, M4))
df$n = unlist(Map(function(l) {c(l$n[[1]][1]) }, M4))
df$type = unlist(Map(function(l) {as.character(l$type[[1]][1])}, M4))
df$h = unlist(Map(function(l) {c(l$h[[1]][1]) }, M4))
df$period = unlist(Map(function(l) {as.character(l$period[[1]][1])}, M4))
```

```{r, message=FALSE, warning=FALSE}
str(df)
table(df$period)
```

## 4. Extract M4 data: yearly, quarterly, monthly, dayly, hourly
```{r, message=FALSE, warning=FALSE}
yearly_M4 = Filter(function(l) l$period == "Yearly", M4)
quarterly_M4 = Filter(function(l) l$period == "Quarterly", M4)
monthly_M4 = Filter(function(l) l$period == "Monthly", M4)
weekly_M4 = Filter(function(l) l$period == "Weekly", M4)
hourly_M4 = Filter(function(l) l$period == "Hourly", M4)
daily_M4 = Filter(function(l) l$period == "Daily", M4)
```

## 5. Plot the first sample 
(including n_ahead data in red)
```{r, message=FALSE, warning=FALSE, echo = FALSE}
plot(ts(c(M4[[1]]$x, M4[[1]]$xx), start = start(M4[[1]]$x), frequency = frequency(M4[[1]]$x)), 
    col = 'black', type = 'l')
lines(M4[[1]]$xx, col = 'red')
```

## 6. Extract one sample month to perform statistical models
Extract the first month: full, training, and test set
```{r, message=FALSE, warning=FALSE}
monthly_1_full = ts(c(monthly_M4[[1]]$x, monthly_M4[[1]]$xx),
               start=start(monthly_M4[[1]]$x), 
               frequency = frequency(monthly_M4[[1]]$x))
monthly_1_train = ts(monthly_M4[[1]]$x, 
                  start=start(monthly_M4[[1]]$x), 
                  frequency = frequency(monthly_M4[[1]]$x))
monthly_1_test = ts(monthly_M4[[1]]$xx, 
                  start=start(monthly_M4[[1]]$xx), 
                  frequency = frequency(monthly_M4[[1]]$xx))
```
Explore the structure of the training and test set
```{r, message=FALSE, warning=FALSE}
head(monthly_1_train)
str(monthly_1_train)
head(monthly_1_test)
str(monthly_1_test)
```
### plot the monthly sample: including full, training and test set
```{r, message=FALSE, warning=FALSE}
plot(ts(monthly_1_full,
        start = start(monthly_1_full),
        frequency = frequency(monthly_1_full)),
     type = 'l', col = 'black', ylab ='', xlab ='')
lines(monthly_1_test, col = 'red', type = 'o', xlab = '')
```
Produce a polar coordinate season plot
```{r, message=FALSE, warning=FALSE}
ggseasonplot(monthly_1_train, polar = T)
```
Create subseries plot that comprises mini time plots for each season
```{r, message=FALSE, warning=FALSE}
ggsubseriesplot(monthly_1_train)
```
### Plot the monthly sample: Removing trend 
```{r, message=FALSE, warning=FALSE}
autoplot(diff(monthly_1_train))
```
## 7. Explore time series patterns: Trend, seasonal, or cyclic
```{r, message=FALSE, warning=FALSE}
acf2(diff(monthly_1_train))
```

## 8. Ljung-Box test: Finding h autocorrelation
```{r, message=FALSE, warning=FALSE}
Box.test(monthly_1_train, fitdf = 0, lag = 24, type = 'Lj')
Box.test(diff(monthly_1_train), fitdf = 0, lag = 24, type = 'Lj')
```
The **Ljung-Box** test result (p-value <<< 0.05) shows autocorrelation in both original data and differencing data.

## 9. Employ statistical model to forecast
### 9.1. Seasonal Naive model
  Model summary and plot
```{r, message=FALSE, warning=FALSE}
snaive_f = snaive(diff(monthly_1_train), h = 18)
autoplot(snaive_f) 
summary(snaive_f)
```
  Fitted values and residuals
```{r, message=FALSE, warning=FALSE}
autoplot(snaive_f, series = "Train")
```
  Check residuals whether white noise or not
```{r, message=FALSE, warning=FALSE}
checkresiduals(snaive_f)
```
  Accuracy of Seasonal Naive model
```{r, message=FALSE, warning=FALSE}
accuracy(snaive_f, diff(monthly_1_test))
```

### 9. Seasonal Arima model
  Model summary and plot
```{r, message=FALSE, warning=FALSE}
sarima_f = auto.arima(monthly_1_train)
summary(sarima_f)
sarima_f %>% 
  forecast(h = 18) %>%
  autoplot()
```
  Check residuals whether white noise or not
```{r, message=FALSE, warning=FALSE}
checkresiduals(sarima_f)
```
  Accuracy of Seasonal Arima model
```{r, message=FALSE, warning=FALSE}
accuracy(forecast(sarima_f, h = 18), monthly_1_test, d = 0, D = 1)
```

